Dominance Regions for Rank Two Cluster Algebras

### 2022-11-04 dominance_regions.tex

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance region} $\cP_\lambda$ (cf. Definition~\ref{def:dominance}).
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance region $\cP_\lambda$ of the $\bfg$-vector $\lambda=(\lambda_0,\lambda_1)$ is the polygon consisting of those $\mu=(\mu_0,\mu_1)\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.2}
      \[
        \begin{array}{rcccl}
          0 & \!\leq\! & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \!\leq\! & \left[-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1\right]_+\!\!;
          \\
          0 & \!\leq\! & -\frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \!\leq\! & \left[c\lambda_0\right]_+;
          \\
          0 & \!\leq\! &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \!\leq\! & \left[\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1\right]_+\!\!;
          \\
          0 & \!\leq\! & -(\mu_0-\lambda_0) + \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \!\leq\! & \left[-b \lambda_1\right]_+;
        \end{array}
      \]
    }
    where $[x]_+:=\max(x,0)$.

  \end{theorem}
  \begin{remark}
    Explicit calculations reveal that for certain pairs of notable bases (e.g. greedy and triangular) most of the coefficients $q_{\lambda,\mu}$ are zero.
    Understanding this phenomenon might be worth further study.
  \end{remark}

  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} (positively) spanned by the vectors $\big(2b,-bc\pm\sqrt{bc(bc-4)}\big)$.
  Lattice points outside of $\cI$ are precisely the $\bfg$-vectors of cluster monomials in $\cA(b,c)$.
  \begin{corollary}
    \label{cor:dominance vertices}
    There are six classes of dominance polygons.
    \begin{enumerate}
      \item If $\lambda$ lies outside of $\cI$, then $\cP_\lambda$ is the point $\lambda$. 
      \item If $\lambda$ lies in the cone spanned by the vectors $(2b,-bc-\sqrt{bc(bc-4)})$ and $(2,-c)$, then $\cP_\lambda$ is the trapezoid with  vertices $\lambda$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, $-\frac{bc+\sqrt{bc(bc-4)}}{2c}\big(\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1,0\big)$, and ${\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(-(bc-2)\lambda_0-b\lambda_1,c\lambda_0+2\lambda_1\big)}$.
      \item If $\lambda$ lies on the ray spanned by $(2,-c)$, then $\cP_\lambda$ is the triangle with vertices $\lambda$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, and $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$.
      \item If $\lambda$ lies in the cone spanned by the vectors $(2,-c)$ and $(b,-2)$, then $\cP_\lambda$ is the kite with  vertices $\lambda$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}(2\lambda_0+b\lambda_1,c\lambda_0+2\lambda_1)$, and $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$.
      \item If $\lambda$ lies on the ray spanned by $(b,-2)$, then $\cP_\lambda$ is the triangle with vertices $\lambda$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, and $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$.
      \item If $\lambda$ lies in the cone spanned by the vectors $(b,-2)$ and $(2b,-bc+\sqrt{bc(bc-4)})$, then $\cP_\lambda$ is the trapezoid with vertices $\lambda$, $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(2\lambda_0+b\lambda_1,-c\lambda_0-(bc-2)\lambda_1\big)$, $-\frac{bc+\sqrt{bc(bc-4)}}{2b}\big(0,\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1\big)$, and $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$.
    \end{enumerate}
  \end{corollary}

@@

>>>>

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We therefore assume that $bc\ge4$.
  Write $\cI \subset \RR^2$ for the \emph{imaginary cone} (positively) spanned by the vectors $\big(2b,-bc\pm\sqrt{bc(bc-4)}\big)$.
  Lattice points outside of $\cI$ are precisely the $\bfg$-vectors of cluster monomials in $\cA(b,c)$.

  We give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance region} $\cP_\lambda$ (cf. Definition~\ref{def:dominance}).
  \begin{theorem}
    \label{th:dominance inequalities}
    If $\lambda$ lies outside of $\cI$, then the dominance region $\cP_\lambda$ is the point $\lambda$.
    Otherwise the dominance region $\cP_\lambda$ of the $\bfg$-vector $\lambda=(\lambda_0,\lambda_1)$ is the polygon consisting of those $\mu=(\mu_0,\mu_1)\in\RR^2$ satisfying the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.2}
      \[
        \begin{array}{rcccl}
          0 & \!\leq\! & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \!\leq\! & -c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1;
          \\
          0 & \!\leq\! & -\frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \!\leq\! & c\lambda_0;
          \\
          0 & \!\leq\! &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \!\leq\! & \frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1;
          \\
          0 & \!\leq\! & -(\mu_0-\lambda_0) + \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \!\leq\! & -b \lambda_1.
        \end{array}
      \]
    }
  \end{theorem}
  \begin{remark}
    Explicit calculations reveal that for certain pairs of notable bases (e.g. greedy and triangular) most of the coefficients $q_{\lambda,\mu}$ are zero.
    Understanding this phenomenon might be worth further study.
  \end{remark}

  \begin{corollary}
    \label{cor:dominance vertices}
    There are six classes of dominance polygons.
    \begin{enumerate}
      \item If $\lambda$ lies outside of $\cI$, then $\cP_\lambda$ is the point $\lambda$. 
      \item If $\lambda$ lies in the cone spanned by the vectors $(2b,-bc-\sqrt{bc(bc-4)})$ and $(2,-c)$, then $\cP_\lambda$ is the trapezoid with  vertices $\lambda$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, $-\frac{bc+\sqrt{bc(bc-4)}}{2c}\big(\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1,0\big)$, and ${\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(-(bc-2)\lambda_0-b\lambda_1,c\lambda_0+2\lambda_1\big)}$.
      \item If $\lambda$ lies on the ray spanned by $(2,-c)$, then $\cP_\lambda$ is the triangle with vertices $\lambda$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, and $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$.
      \item If $\lambda$ lies in the cone spanned by the vectors $(2,-c)$ and $(b,-2)$, then $\cP_\lambda$ is the kite with  vertices $\lambda$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}(2\lambda_0+b\lambda_1,c\lambda_0+2\lambda_1)$, and $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$.
      \item If $\lambda$ lies on the ray spanned by $(b,-2)$, then $\cP_\lambda$ is the triangle with vertices $\lambda$, $\big(0,\frac{bc+\sqrt{bc(bc-4)}}{2b}\lambda_0+\lambda_1\big)$, and $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$.
      \item If $\lambda$ lies in the cone spanned by the vectors $(b,-2)$ and $(2b,-bc+\sqrt{bc(bc-4)})$, then $\cP_\lambda$ is the trapezoid with vertices $\lambda$, $\frac{bc+\sqrt{bc(bc-4)}}{2\sqrt{bc(bc-4)}}\big(2\lambda_0+b\lambda_1,-c\lambda_0-(bc-2)\lambda_1\big)$, $-\frac{bc+\sqrt{bc(bc-4)}}{2b}\big(0,\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1\big)$, and $\big(\lambda_0+\frac{bc+\sqrt{bc(bc-4)}}{2c}\lambda_1,0\big)$.
    \end{enumerate}
  \end{corollary}

@@

  \begin{tikzpicture}[scale=.45]
